---
title: 'Run 9: Oman 2018 NO$_3$ $\delta^{15}N$ and $\delta^{18}O$'
subtitle: "Source file: 20180727_run9_calibration.Rmd"
date: "`r Sys.Date()`"
output:
  html_document: 
    css: stylesheet.css
    fig_caption: yes
    number_sections: yes
    df_print: paged
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(isoreader)
library(isorunN2O)
library(isoprocessor)
library(stringr)
library(readxl)
library(latex2exp)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(readxl)
library(plotly)
library(knitr)
opts_knit$set(root.dir = ".", output_dir = file.path("documents"))
opts_chunk$set(
  dev=c("png", "pdf"), dev.args=list(pdf = list(encoding="WinAnsi", useDingbats=FALSE)),
  fig.keep="all")
source(file.path("scripts", "functions.R"))
```


The dissolved NO$_3$ isotopic composition was measured after nitrite removal using the denitrifier method coupled to GC-IRMS as described in the methods section of the manuscript. The following code steps through the calibration of measured isotopic values. The fluid samples processed here in this document were collected in Oman in the February 2018 field season.


This processing script uses isoreader version `r packageVersion("isoreader")`, isorunN2O version `r packageVersion("isorunN2O")`, and isoprocessor version `r packageVersion("isoprocessor")`.


## Load Raw Data
### To create .rds file from folder of .dxf files for fast loading
```{r eval = FALSE}
#Load raw data from .dxf files
  folder <- "180727_OM_run9"
  raw_data <-
    file.path("data", "raw", folder) %>%
    iso_read_continuous_flow()
  #look for problematic files
 if (iso_has_problems(raw_data)) {
   message("are these problems okay?")
    print(iso_get_problems_summary(raw_data))
  }

#omit any files with problems and export to an .rds file
  raw_data %>%
    iso_omit_files_with_problems() %>%
    iso_save("data/raw/run9_raw") 
```

### Load raw data from .rds file
```{r "load run"}
# load entire run
raw_data <- iso_read_continuous_flow(file.path("data","raw", "run9_raw.cf.rds"))
```

### Parse file information
```{r "retrieve data"}
df.raw <- raw_data %>% 
  # aggregate peak data
  iso_get_vendor_data_table(include_file_info = c(file_path, file_datetime, Row, `Identifier 1`, `Identifier 2`, Analysis)) %>% 
  # select N2O peak (select_N2O_peak requires file field)
  rename(file = file_id) %>% 
  select_N2O_peak(c(310, 370)) %>% 
  # extract relevant file information
  mutate(folder = basename(dirname(file_path)), run_number = parse_integer(Row), analysis = parse_integer(Analysis),
         name = `Identifier 1`, volume = parse_number(`Identifier 2`)) %>% 
  # select relevant columns
  select(analysis, 
         folder, date = file_datetime, run_number, volume,
         area = `Intensity All`, d45 = `d 45N2O/44N2O`, d46 = `d 46N2O/44N2O`) %>% 
  # pull in categories and names from metadata file by joining with analysis # column
  left_join(read_excel(file.path("data", "metadata", "20180727_category_metadata.xlsx")) %>% select(-expt), by = c("analysis")) %>% 
  # convert volumes from uL to mL 
  mutate(
    volume = volume/1000,
    #add big picture categories
    is_standard = str_detect(name, "(USGS|IAEA)"),
    is_sample = category %in% c("Oman")
  ) %>% #select only relevant categories from metadata for downstream processing
  filter(category %in% c("Background", "Control", "N2O", "HPW", "Conditioner", "Oman", "USGS-34 NR", "IAEA-NO3 NR")) %>%
	select(-timepoint, -CvsK) %>%
	#some standards were subjected to nitrite removal using 5% sulfamic; exclude these from calibration by changing category to excluding and changing is_standard to FALSE
	mutate(category = ifelse(str_detect(name, "5%"), "excluded", category),
				 is_standard = ifelse(str_detect(name, "5%"), FALSE, is_standard))

df.raw
```


### Summary of the raw data
$\delta^{45}N_{2}O$ and $\delta^{46}N_{2}O$ in $‰$ are represented as d45 and d46 respectively in the data table. Raw d45 and d46 peak areas and standard deviations are reported. Two potassium nitrate isotopic standards (USGS-34 and IAEA) dissolved in HPW (high purity water) in two concentrations (30 $\mu$M and 250 $\mu$M) were subjected to nitrite removal protocols and spaced every 8-10 samples for calibration of  $\delta^{15}N$, $\delta^{18}O$, and nitrate concentration. Vials of 20 nmoles of injected N2O were also evenly spaced every 8 samples + standards to monitor drift.
```{r "raw data summary"}
#table of average and standard deviations of peak areas, d45, and d46 for standards and controls
df.raw %>% group_by(category, name) %>% 
  iso_summarize_data_table(area, d45, d46, cutoff = 3)
df.raw
```


## First look at standards
With the exception of run_number 88, standards look fairly uniform in area and d45.
```{r "plot standards", warning= FALSE, plotly=TRUE}
plot_raw_standards <- df.raw %>%
	#select just isotopic and N2O standards
	filter(category %in% c("IAEA-NO3 NR", "USGS-34 NR", "N2O")) %>%
  #plot with panels with ggplot  
  iso_plot_data(
  x = run_number, y = d45, points = TRUE,
  # shape = 21 with fill makes data points with black borders
  shape = 21, fill = name, size = area,
  panel = category ~ .
)

#change to colorblind friendly palette and add axis labels
plot_raw_standards + 
  	scale_fill_manual(values = cbPalette) + 
	  scale_color_manual(values = cbPalette) +
    ylab(latex2exp::TeX("$\\delta^{45}N_{2}O$")) +
    xlab("run number") 

#plot interactive version
ggplotly(plot_raw_standards, dynamicTicks = TRUE)
```


### Identify problematic data points
```{r "remove problematic data points"}
 df.raw <- df.raw %>% 
 change_category(run_number == 88, "excluded") #area too small
```

## Corrections

### Drift correction
Evaluate drift throughout run duration. Slight drift is observed; linear correction applied.
```{r}
df.drift <- df.raw %>% evaluate_drift(d45, d46, correct = TRUE, method = "lm", correct_with = category %in% c("USGS-34 NR", "IAEA-NO3 NR", "N2O"), plot = TRUE)
#linear calibration applied; new columns d45.drift, d46.drift, p.drift created
```


### $^{17}$O correction
Correction for $^{17}$O isobaric interference in the mass 45 peak according to Kaiser et al. (2008) to derive raw $\delta^{15}N$ and $\delta^{18}O$ values.
```{r}
df.O17 <- df.drift %>%
  correct_N2O_for_17O(d45.drift, d46.drift) %>% #apply the 17O corrections to the drift-corrected d45.drift and d46.drift columns
	#corrected columns are d15.raw and 18.raw; p.17Ocor column added with correction parameters
  select_columns(-d45, -d45.drift, -d46, -d46.drift) 
```


## Concentration Calibration with Isoprocessor

### Load Standards
```{r}
#load in concentration and isotopic standards from excel file
stds_isotopes <- read_excel(file.path("data", "standards", "standards_run9.xlsx"), sheet = "isotopes")
stds_concs <- read_excel(file.path("data", "standards", "standards_run9.xlsx"), sheet = "concentrations")
stds_isotopes
stds_concs
```

### Generate concentration calibration
```{r "add standards", warning = FALSE}
df_calibp <- 
  df.O17 %>% 
  filter(category != "excluded") %>% #filter out excluded categories and conditioners
  filter(category != "Conditioner") %>%
  select(-p.17Ocor, -p.drift, -date) %>% #removed nested parameter columns
  # add standards
  iso_add_standards(stds_isotopes, match_by = category, is_std_peak = is_iso_std) %>% #add isotope standards
  iso_add_standards(stds_concs, match_by = name, is_std_peak = is_conc_std)  #add concentration standards
```

```{r "concentration calibration"}
df_calib <- df_calibp %>% 
	 mutate(amount.nmol = true_conc.uM * volume / dilution) %>% #calculate nmoles from concentration
  iso_prepare_for_calibration() %>% 
  iso_generate_calibration(
    model = lm(area ~ amount.nmol - 1), #linear model where area is the response variable; the - 1 removes the implied intercept term
    calibration = "area",
    use_in_calib = is_conc_std) 

# check for problems
df_calib %>% iso_get_problematic_calibrations(calibration = "area")
```

### Inspect calibration
Residuals ~5% for most concentration standards
```{r warning = FALSE}
# parameter overview table
df_calib %>% 
  iso_get_calibration_parameters(calibration = "area")
```

```{r warning=FALSE}
# visualization of residuals
conc_resid <- df_calib %>% 
  iso_get_calibration_data() %>% 
  filter(is_conc_std) %>%
	filter(category != "excluded") %>% #filter for only standards used in calibration
  mutate(
    `Var: rel. residual area [%]` = 100*area_resid/area #calculate % relative residual area
  ) %>%
	ggplot(
		aes(x = run_number, y =`Var: rel. residual area [%]`, color = name)) +
			geom_point(size = 4) +
			theme_figure() +
	scale_color_manual(values = cbPalette) +
			xlab("Run Number") +
			ylab("Relative Residual Area (%)") + 
	guides(color = guide_legend(title = NULL)) 
	
ggplotly(conc_resid) 
```

### Apply calibration
```{r warning = FALSE}
# apply concentration calibration
df_w_conc <- df_calib %>% 
  iso_apply_calibration(
    predict = amount.nmol,
    calibration = "area",
    calculate_error = TRUE, #standard error using Wald method
    predict_range = c(0, 20) #prediction range from standards
  ) %>% 
  iso_get_calibration_data()

df_w_conc
```

### Check data
A concentration control was not used in calibration, but is calculated to be 167.6 uM +/- 1.2 uM. The known value is 161 uM
```{r}
# visualize (note that most are technically out of range of the calibration which is only defined for the narrow signal observed in standards; the rest are extrapolations beyond the calibration model)
plot_nmoles <- df_w_conc %>%
	filter(category != "excluded") %>%
	ggplot(
	#define global plot aesthetics 
		aes(x = run_number, y = amount.nmol_pred, color = category, label = name)) +
	scale_color_manual(values = cbPalette) + 
	xlab("Run number") +
	ylab("Predicted nmoles of N") +
	theme_figure(grid = TRUE, legend = TRUE, text_size = 20) +
	guides(color = guide_legend(title = NULL))

#ggplot version with check if calibration is ok
plot_nmoles_ggplot <- plot_nmoles + 
	geom_point(aes(shape = df_w_conc$area_calib_ok, fill = df_w_conc$category, color = df_w_conc$category)) + 
	scale_shape_manual(values = c(21, 22)) + 
	scale_fill_manual(values = cbPalette) + 
	theme_figure(grid = TRUE, legend = TRUE, text_size = 20) + 
	guides(shape = guide_legend(title = "Calibration OK"), fill = guide_legend(title = NULL))
plot_nmoles_ggplot

#interactive plot to see individual sample names from plot above	
ggplotly((plot_nmoles + geom_point()), tooltip=c("name", "amount.nmol_pred", "run_number"))
```

```{r "exclude samples with <5 nmoles N"}
#exclude samples that have <5 nmoles predicted N
df_w_conc <- df_w_conc %>%
	filter(
		if_else(is_sample == TRUE, amount.nmol_pred > 5, amount.nmol_pred > 0)
	)
df_w_conc
```

```{r "calculate concentration"}
# calculate concentration
df_w_conc <- 
  df_w_conc %>% 
  mutate(conc_uM = dilution * amount.nmol_pred / volume,
         conc_uM_err = dilution * amount.nmol_pred_se / volume)

# check amount and concentration data
df_w_conc %>% filter(category != "N2O") %>%
  group_by(category, true_conc.uM) %>% 
  iso_summarize_data_table(
    conc_uM,
    amount.nmol_pred, 
    amount.nmol_pred_se, 
    cutoff = 2) 
```


## Delta calibration with Isoprocessor

### Generate calibration for N
```{r warning = FALSE}
df_calibN <- df_w_conc %>% 
  iso_prepare_for_calibration() %>% 
  iso_generate_calibration(
    model = c(
      # generate calibration using different regression models 
      Nlinear = lm(d15.raw ~ true_d15N), #simple linear
      Nwith_conc = lm(d15.raw ~ true_d15N + conc_uM), #multiple regression with concentration
      Nwith_volume = lm(d15.raw ~ true_d15N + volume) #multiple regression with volume
    ),
    calibration = "d15N",
    use_in_calib = is_iso_std) 
```

### Inspect calibration for N
```{r}
# parameter overview table
df_calibN %>% 
  iso_get_calibration_parameters(calibration = "d15N")

#visualization of residuals for each N isotope calibration
N_residuals <- df_calibN %>%
	iso_get_calibration_data() %>%
	filter(is_iso_std) %>%
	filter(category != "excluded") %>% #filter for only standards used in calibration
	ggplot(
		aes(x = run_number, y = d15N_resid, color = d15N_calib)) +
			geom_point(size = 4) +
			theme_figure() +
	scale_color_manual(values = cbPalette) +
			xlab("Run Number") +
			ylab("Residuals") + 
	guides(color = guide_legend(title = NULL)) 
	
ggplotly(N_residuals) 

#applying simple linear regression as more complicated multiple regressions yield very similar results
```

### Generate calibration for O
```{r warning = FALSE}
df_calibO <- df_w_conc %>% 
  iso_prepare_for_calibration() %>% 
  iso_generate_calibration(
    model = c(
      # generate calibration using different regression models
      Olinear = lm(d18.raw ~ true_d18O), #linear
      Owith_conc = lm(d18.raw ~ true_d18O + conc_uM), #multiple regression with concentration
      Owith_volume = lm(d18.raw ~ true_d18O + volume) #multiple regression with volume
    ),
    calibration = "d18O",
    use_in_calib = is_iso_std) 
```

### Inspect calibration for O
```{r}
# parameter overview table
df_calibO %>% 
  iso_get_calibration_parameters(calibration = "d18O")
  
#visualization of residuals for each N isotope calibration
O_residuals <- df_calibO %>%
	iso_get_calibration_data() %>%
	filter(is_iso_std) %>%
	filter(category != "excluded") %>% #filter for only standards used in calibration
	ggplot(
		aes(x = run_number, y = d18O_resid, color = d18O_calib)) +
			geom_point(size = 4) +
			theme_figure() +
	scale_color_manual(values = cbPalette) +
			xlab("Run Number") +
			ylab("Residuals") + 
	guides(color = guide_legend(title = NULL)) 
	
ggplotly(O_residuals) 

#multiple regressions using concentration or volume are better than the simple linear calibration model# visualization of residuals
```

### Apply calibration
Applying a multiple regression calibration with volume for N and O
```{r warning = FALSE}
# apply linear calibration for N
df_finalN <- df_calibN %>%
filter(d15N_calib == "Nlinear")%>% 
  iso_apply_calibration(
    predict = true_d15N,
    calibration = "d15N",
    calculate_error = TRUE
  ) %>%
  iso_get_calibration_data() %>% iso_remove_list_columns()
 
# apply multiple regression calibration with volume for O
df_finalO <- df_calibO %>% 
  filter(d18O_calib == "Owith_volume") %>% 
  iso_apply_calibration(
    predict = true_d18O,
    calibration = "d18O",
    calculate_error = TRUE
  ) %>%
 iso_get_calibration_data() %>%
  iso_remove_list_columns()

#join O and N calibrated data into final dataframe, df_final
 df_final <- left_join(df_finalN, df_finalO)
 
 df_final
```

## Check data
```{r}
# summary of calibrated isotopic data for major categories
df_final %>% 
  group_by(category, true_d15N, true_d18O) %>% 
  iso_generate_summary_table(
    d15N = true_d15N_pred, 
    d15N_se = true_d15N_pred_se,
    d18O = true_d18O_pred, 
    d18O_se = true_d18O_pred_se,
    cutoff = 3)
```

```{r "Visualize Calibrated Nitrogen Isotopic Data", plotly = TRUE}
#visualization of calibrated N isotopic data
N_plot <- df_final %>%
	filter(category != "excluded") %>% #remove excluded samples and standards from plot
	ggplot(
		aes(x = run_number, y = true_d15N_pred, color = category, label = name)) +
			geom_point(size = 4) +
		  geom_errorbar(aes(ymax = true_d15N_pred + true_d15N_pred_se, ymin = true_d15N_pred - true_d15N_pred_se)) +
			theme_figure() +
	scale_color_manual(values = cbPalette) +
			xlab("Run Number")  +
	guides(color = guide_legend(title = NULL)) 

#ggplot version with check if calibration is ok 
plot_Nisotopes_ggplot <- N_plot + 
	geom_point(aes(shape = df_final$d15N_calib_ok, fill = df_final$category, color = df_final$category)) + 
	scale_fill_manual(values = cbPalette) + 
	theme_figure(grid = TRUE, legend = TRUE, text_size = 20) + 
	ylab(latex2exp::TeX("$\\delta^{15}N$")) +
	guides(shape = guide_legend(title = "Calibration OK"), fill = guide_legend(title = NULL), color = guide_legend(title = NULL))

plot_Nisotopes_ggplot

#plot interactive version to see individual sample names from plot above
N_plot_interactive <- ggplotly(N_plot + theme_bw()) %>% 
	layout(xaxis = list(title = plotly::TeX("Run  Number")), yaxis = list(title = plotly::TeX("\\delta^{15}N"))) %>% 
	config(mathjax = 'cdn')

N_plot_interactive
```


```{r "Visualize Calibrated Oxygen Isotopic Data", plotly = TRUE}
#visualization of calibrated O isotopic data
O_plot <- df_final %>%
	filter(category != "excluded") %>% #remove excluded samples and standards from plot
	ggplot(
		aes(x = run_number, y = true_d18O_pred, color = category, label = name)) +
			geom_point(size = 4) +
		  geom_errorbar(aes(ymax = true_d18O_pred + true_d18O_pred_se, ymin = true_d18O_pred - true_d18O_pred_se)) +
			theme_figure() +
	scale_color_manual(values = cbPalette) +
			xlab("Run Number")  +
	guides(color = guide_legend(title = NULL)) 

#ggplot version with check if calibration is ok
plot_Oisotopes_ggplot <- O_plot + 
	geom_point(aes(shape = df_final$d18O_calib_ok, fill = df_final$category, color = df_final$category)) + 
	scale_fill_manual(values = cbPalette) + 
	theme_figure(grid = TRUE, legend = TRUE, text_size = 20) + 
	ylab(latex2exp::TeX("$\\delta^{18}O$")) +
	guides(shape = guide_legend(title = "Calibration OK"), fill = guide_legend(title = NULL), color = guide_legend(title = NULL))

plot_Oisotopes_ggplot

#plot interactive version to see individual sample names from plot above
O_plot_interactive <- ggplotly(O_plot + theme_bw()) %>% 
	layout(xaxis = list(title = plotly::TeX("Run  Number")), yaxis = list(title = plotly::TeX("\\delta^{18}O"))) %>% 
	config(mathjax = 'cdn')

O_plot_interactive
```


## Summary
```{r}
df_final <- df_final %>%
  mutate(
    conc = conc_uM,
    conc_err = conc_uM_err,
    d15N = true_d15N_pred,
    d15N_err = true_d15N_pred_se,
    d18O = true_d18O_pred,
    d18O_err = true_d18O_pred_se,
    compound = "NO3"
  )  

df_final %>% group_by(category, name) %>% 
  iso_summarize_data_table(cutoff = 1, d15N, d15N_err, d18O, d18O_err, conc, conc_err)
```


## Export sample data
```{r, include=FALSE}
df_final %>%
  filter(category == "Oman") %>%
  select(analysis, run_number, volume, area, pH, name, category, dilution, col_NO3, conc, conc_err, d15N, d15N_err, d18O, d18O_err, compound) %>%
write.csv(file = file.path("data", "calibrated", "run9_calib_data_export.csv"))

df_final %>% 
 filter(category == "Oman") %>% 
  saveRDS(file = file.path("data", "calibrated", "df_final_run9"))
# saveRDS
``` 


## Export standards data
```{r, include=FALSE}
df_final %>%
  filter(category %in% c("USGS-34", "IAEA-NO3")) %>%
  select(analysis, run_number, volume, area, name, category, conc, conc_err, d15N, d15N_err, d18O, d18O_err, compound) %>%
write.csv(file = file.path("data", "calibrated", "run9_stds_calib_data_export.csv"))

df_final %>% 
 filter(category %in% c("USGS-34", "IAEA-NO3")) %>%
  saveRDS(file = file.path("data", "calibrated", "df_stds_run9"))
# saveRDS
```
