---
title: 'Run 10: Oman 2018 Organic N $\delta^{15}N$'
subtitle: "Source file: 20180731_run10_calibration.Rmd"
date: "`r Sys.Date()`"
output:
  html_document: 
    css: stylesheet.css
    fig_caption: yes
    number_sections: yes
    df_print: paged
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(isoreader)
library(isorunN2O)
library(isoprocessor)
library(stringr)
library(readxl)
library(latex2exp)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(readxl)
library(plotly)
library(knitr)
opts_knit$set(root.dir = ".", output_dir = file.path("documents"))
opts_chunk$set(
  dev=c("png", "pdf"), dev.args=list(pdf = list(encoding="WinAnsi", useDingbats=FALSE)),
  fig.keep="all")
source(file.path("scripts", "functions.R"))
```


Organic N isotopic composition was measured using the denitrifier method coupled to GC-IRMS after a persulfate oxidation as described in the methods section of the manuscript. The following code steps through the calibration of measured isotopic values. The fluid samples processed here in this document were collected in Oman in the February 2018 field season.


This processing script uses isoreader version `r packageVersion("isoreader")`, isorunN2O version `r packageVersion("isorunN2O")`, and isoprocessor version `r packageVersion("isoprocessor")`.


## Load Raw Data
### To create .rds file from folder of .dxf files for fast loading
```{r eval = FALSE}
#Load raw data from .dxf files
  folder <- "180731_OM_run10"
  raw_data <-
    file.path("data", "raw", folder) %>%
    iso_read_continuous_flow()
  #look for problematic files
 if (iso_has_problems(raw_data)) {
   message("are these problems okay?")
    print(iso_get_problems_summary(raw_data))
  }

#omit any files with problems and export to an .rds file
  raw_data %>%
    iso_omit_files_with_problems() %>%
    iso_save("data/raw/run10_raw") 
```

### Load raw data from .rds file
```{r "load run"}
#load entire run
raw_data <- iso_read_continuous_flow(file.path("data", "raw", "run10_raw.cf.rds"))
```

### Parse file information
```{r "retrieve data"}
df.raw <- raw_data %>% 
  # aggregate peak data
  iso_get_vendor_data_table(include_file_info = c(file_path, file_datetime, Row, `Identifier 1`, `Identifier 2`, Analysis)) %>% 
  # select N2O peak (select_N2O_peak requires file field)
  rename(file = file_id) %>% 
  select_N2O_peak(c(310, 370)) %>% 
  # extract relevant file information
  mutate(folder = basename(dirname(file_path)), run_number = parse_integer(Row), analysis = parse_integer(Analysis),
         name = `Identifier 1`, volume = parse_number(`Identifier 2`)) %>% 
  # select relevant columns
  select(analysis,  
         folder, date = file_datetime, run_number, volume,
         area = `Intensity All`, d45 = `d 45N2O/44N2O`, d46 = `d 46N2O/44N2O`) %>% 
	 # pull in categories from metadata file instead of extracting from name
  left_join(read_excel(file.path("data", "metadata", "180731_category_metadata.xlsx")) %>% select(-expt), by = c("analysis")) %>% 
  # big picture categories
  mutate(
    is_standard = str_detect(name, "(USGS|IAEA)"),
    is_sample = category == "Oman",
    volume = volume/1000
  ) %>% 
#select only relevant categories from metadata for downstream processing
  filter(category %in% c("Savillex", "Oxidation blank", "control", "N2O", "Conditioner", "Oman", "USGS-34", "IAEA-NO3", "USGS-40", "USGS-65"))

df.raw
``` 

```{r "Calculate backgrounds"}
df.raw <- df.raw %>%
 # backgrounds (these are NOT used in the calibration calculations but are good for sanity checks)
  calculate_background(area, crit = grepl("Savillex", name)) %>% 
  calculate_oxidation_blank(area, volume, crit = grepl("Oxidation blank", name))  
 
df.raw
```


## Summary of the raw data
$\delta^{45}N_{2}O$ in $‰$ is represented as d45 in the data table. Raw d45 peak areas and standard deviations are reported. Two potassium nitrate isotopic standards (USGS-34 and IAEA) dissolved in HPW (high purity water) at a 37 $\mu$M concentration were spaced every 8-10 samples for calibration of  $\delta^{15}N$ of nitrate. Vials of 20 nmoles of injected N2O were also evenly spaced every 8 samples + standards to monitor drift. In addition to nitrate isotopic standards, organic N standards, USGS-40 and USGS-65 were oxidized with persulfate along with samples to constrain oxidation efficiency and constrain error associated with the oxidation protocol.

```{r "raw data summary"}
#table of average and standard deviations of peak areas, d45, and d46 for standards and controls
df.raw %>% group_by(category, name) %>% 
  iso_summarize_data_table(area, d45, cutoff = 2) 
```

## First look at standards
Note: Injection volumes were half the intended 20 nmoles.
```{r "plot standards", warning= FALSE, plotly=TRUE}
plot_raw_standards <- df.raw %>%
	#select just isotopic and N2O standards
	filter(category %in% c("IAEA-NO3", "USGS-34", "N2O", "USGS-65", "USGS-40")) %>%
  #plot with panels with ggplot  
  iso_plot_data(
  x = run_number, y = d45, points = TRUE,
  # shape = 21 with fill makes data points with black borders
  shape = 21, fill = name, size = area,
  panel = category ~ .
)

#change to colorblind friendly palette and add axis labels
plot_raw_standards + 
  	scale_fill_manual(values = cbPalette) + 
	  scale_color_manual(values = cbPalette) +
    ylab(latex2exp::TeX("$\\delta^{45}N_{2}O$")) +
    xlab("run number") 

#plot interactive version
ggplotly(plot_raw_standards, dynamicTicks = TRUE)
```

### Identify problematic data points
```{r "remove problematic data points"}
 df.raw <- df.raw %>% 
 change_category(run_number %in% c(11, 48, 87), "excluded") # 48 and 87 are too large; 11 too small 
```


## Corrections

### Drift correction
Evaluate drift throughout run duration. Only minor drift is observed; linear correction applied.
```{r}
df.drift <- df.raw %>% evaluate_drift(d45, d46, correct = TRUE, method = "lm", correct_with = category %in% c("USGS-34", "IAEA-NO3", "N2O"), plot = TRUE)
#linear calibration applied; new columns d45.drift, d46.drift, p.drift created
```

### $^{17}$O correction
Correction for $^{17}$O isobaric interference in the mass 45 peak according to Kaiser et al. (2008) to derive raw $\delta^{15}N$ and $\delta^{18}O$ values.
```{r}
df.O17 <- df.drift %>%
  correct_N2O_for_17O(d45.drift, d46.drift) %>% #apply the 17O corrections to the drift-corrected d45.drift and d46.drift columns
	#corrected columns are d15.raw and 18.raw; p.17Ocor column added with correction parameters
  select_columns(-d45, -d45.drift, -d46, -d46.drift) 
```


## Concentration Calibration with Isoprocessor

### Load Standards
```{r}
#load in concentration and isotopic standards from excel file
stds_concs <- read_excel(file.path("data", "standards", "standards_run10.xlsx"), sheet = "concentrations")
stds_concs
```

### Generate concentration calibration
```{r "add standards", warning = FALSE}
df_calibp <- 
  df.O17 %>% 
  filter(category != "excluded") %>% #filter out excluded categories and conditioners
  filter(category != "Conditioner") %>%
  select(-p.17Ocor, -p.drift, -date) %>% #removed nested parameter columns
  # add standards
  iso_add_standards(stds_concs, match_by = name, is_std_peak = is_conc_std)  #add concentration standards
```

```{r "concentration calibration"}
df_calib <- df_calibp %>% 
	 mutate(amount.nmol = true_conc.uM * volume / dilution) %>% #calculate nmoles from concentration
  iso_prepare_for_calibration() %>% 
  iso_generate_calibration(
    model = lm(area ~ amount.nmol - 1), #linear model where area is the response variable; the - 1 removes the implied intercept term
    calibration = "area",
    use_in_calib = is_conc_std) 

# check for problems
df_calib %>% iso_get_problematic_calibrations(calibration = "area")
```

## Inspect calibration
Residuals ~10% when using a calibration that includes N org standards; using just nitrate standards this is only 2% 
```{r warning = FALSE}
# parameter overview table
df_calib %>% 
  iso_unnest_calibration_parameters(calibration = "area")
```

```{r warning=FALSE}
# visualization of residuals
conc_resid <- df_calib %>% 
  iso_get_calibration_data() %>% 
  filter(is_conc_std) %>%
	filter(category != "excluded") %>% #filter for only standards used in calibration
  mutate(
    `Var: rel. residual area [%]` = 100*area_resid/area #calculate % relative residual area
  ) %>%
	ggplot(
		aes(x = run_number, y =`Var: rel. residual area [%]`, color = name)) +
			geom_point(size = 4) +
			theme_figure() +
	scale_color_manual(values = cbPalette) +
			xlab("Run Number") +
			ylab("Relative Residual Area (%)") + 
	guides(color = guide_legend(title = NULL)) 
	
ggplotly(conc_resid) 
```

## Apply calibration

```{r warning = FALSE}
# apply concentration calibration
df_w_conc <- df_calib %>% 
  iso_apply_calibration(
    predict = amount.nmol,
    calibration = "area",
    calculate_error = TRUE, #standard error using Wald method
    predict_range = c(0, 20) #prediction range from standards
  ) %>% 
  iso_get_calibration_data()
df_w_conc
```

### Check data
```{r}
# visualize (note that most are technically out of range of the calibration which is only defined for the narrow signal observed in standards; the rest are extrapolations beyond the calibration model)
plot_nmoles <- df_w_conc %>%
	filter(category != "excluded") %>%
	ggplot(
	#define global plot aesthetics 
		aes(x = run_number, y = amount.nmol_pred, color = category, label = name)) +
	scale_color_manual(values = cbPalette) + 
	xlab("Run number") +
	ylab("Predicted nmoles of N") +
	theme_figure(grid = TRUE, legend = TRUE, text_size = 20) +
	guides(color = guide_legend(title = NULL))

#ggplot version with check if calibration is ok
plot_nmoles_ggplot <- plot_nmoles + 
	geom_point(aes(shape = df_w_conc$area_calib_ok, fill = df_w_conc$category, color = df_w_conc$category)) + 
	scale_shape_manual(values = c(21, 22)) + 
	scale_fill_manual(values = cbPalette) + 
	theme_figure(grid = TRUE, legend = TRUE, text_size = 20) + 
	guides(shape = guide_legend(title = "Calibration OK"), fill = guide_legend(title = NULL))
plot_nmoles_ggplot

#interactive plot to see individual sample names from plot above	
ggplotly((plot_nmoles + geom_point()), tooltip=c("name", "amount.nmol_pred", "run_number")) 
```

```{r "exclude samples with <5 nmoles N"}
#exclude samples that have <5 nmoles predicted N
df_w_conc <- df_w_conc %>%
	filter(
		if_else(is_sample == TRUE, amount.nmol_pred > 5, amount.nmol_pred > 0)
	)
df_w_conc
```

```{r "calculate concentration"}
# calculate concentration
df_w_conc <- 
  df_w_conc %>% 
  mutate(conc_uM = dilution * amount.nmol_pred / volume,
         conc_uM_err = dilution * amount.nmol_pred_se / volume)

# check amount and concentration data
df_w_conc %>% filter(category != "N2O") %>%
  group_by(category, true_conc.uM) %>% 
  iso_generate_summary_table(
    conc_uM,
    amount.nmol_pred, 
    amount.nmol_pred_se, 
    cutoff = 2) 
```

### Yields
Samples were all less than ~30 uM after dilution with persulfate solution, so are well constrained by standards.
```{r}
#calculate yields
 df_w_conc <- df_w_conc %>% mutate(
    colorimetric_conc = col_NOx + col_NH4, #persulfate oxidation oxidizes everything to nitrate, so the measurement is for any organic N + any oxidized speices already in the sample
    yield = ifelse(is_standard, conc_uM/true_conc.uM, conc_uM/colorimetric_conc)
  )


p <- df_w_conc %>% 
  filter(category %in% c("IAEA-NO3", "USGS-34", "USGS-40", "USGS-65")) %>% 
  ggplot() +
  aes(x = conc_uM, y = true_conc.uM, text = name, color = category) +
  geom_point(size=2) +
	scale_color_manual(values = cbPalette) +
  geom_abline()  +
  theme_figure() +
	ylab("True concentration [uM]") +
	xlab("Measured concentration [uM]") +
	guides(color = guide_legend(title = NULL))
ggplotly(p) 
```


## Delta Calibration with isorunN2O

### Compare one and two step calibrations 
```{r}
# one step calibration
one_step <- df_w_conc %>% 
  # run organic d15 calibration all in one starting from raw d15
  calibrate_d15_org(d15.raw, area, volume, standards = c("USGS-40" = -4.50, "USGS-65" = 20.68)) %>% 
  mutate(method = "one step") %>% 
  ungroup()

# two step calibration
two_step <- df_w_conc %>% 
  # run regular d15 calibration
  calibrate_d15(
    # regular denitrifier calibration starting from raw d15
    d15.raw, area, standards = c("USGS-34" = -1.8, "IAEA-NO3" = 4.7)) %>%
  # followed by organic d15 calibration
  calibrate_d15_org(
    # now use the already nitrate calibrated d15 as starting point
    d15.cal, area, volume, standards = c("USGS-40" = -4.50, "USGS-65" = 20.68),
    infer_ref_gas = FALSE, infer_bgrd = FALSE) %>%
  mutate(method = "two step") %>% 
  ungroup()
```
## Check standards

```{r}
bind_rows(one_step, two_step) %>% 
  filter(grepl("USGS-(40|65)",category)) %>% 
  group_by(method, category) %>% 
  iso_summarize_data_table(d15.raw, d15.cal, d15.ocal) 
```


### Compare parameters 
Compare the model fit parameters between direct estimation, one-step and two-step methods.
```{r}
params <- bind_rows(one_step, two_step) %>% 
  select(method, starts_with("p.")) %>%
  rename(dir_p.bgrd_area = p.bgrd, dir_p.oblank_ratio = p.oblank) %>% 
  # calculate direct estimate of the oblank_d15
  mutate(
    dir_p.oblank_d15 = (p.oblank_ratio * p.oblank_d15) / dir_p.oblank_ratio,
    p.beta4.err =  sqrt( (p.oblank_d15.err/p.oblank_d15)^2 - (p.oblank_ratio.err/p.oblank_ratio)^2 ),
    dir_p.oblank_d15.err = p.beta4.err/dir_p.oblank_ratio) %>% 
  select(-p.beta4.err) %>% 
  unique() %>% 
  tidyr::gather(param, value, -method) %>% 
  mutate(value = suppressWarnings(as.numeric(value))) %>% 
  filter(!is.na(value))
  
params <- left_join(
  params %>% filter(!grepl(".err", param)),
  params %>% filter(grepl(".err", param)) %>% 
    mutate(param = sub(".err", "", param)) %>% 
    rename(error = value),
  by = c("method", "param")) %>% 
  arrange(param, method) %>% 
  mutate(
    method = ifelse(grepl("dir_", param), "direct", method),
    param = sub("(dir_)?p.", "", param)) %>% 
  unique()
params

params %>% 
  ggplot() +
  aes(method, value, fill = method) + 
  geom_errorbar(aes(ymin = value - 2* error, ymax = value + 2* error)) +
  geom_point(shape = 21, size = 5) +
  labs(x = "", y = "error") + theme_bw()
```

## Compare the calibrations
Looking at the difference between the two calibrations for both batches- except for background calculations, not a major difference 
```{r single_data_plots, fig.width = 10, fig.height = 9}
all_data <-
  left_join(
    two_step %>% rename(d15.ocal_2steps = d15.ocal), 
    one_step %>% select(run_number, name, d15.ocal_1step = d15.ocal), 
    by = c("name", "run_number")) %>% 
  mutate(step_diff = d15.ocal_2steps - d15.ocal_1step) %>% 
  arrange(run_number) %>% 
  filter(category != "excluded") %>% 
  ungroup()

  (all_data %>%
    ggplot() +
    aes(area, abs(step_diff), text = run_number, fill = category) +
    geom_point(size = 4, shape = 21) +
  		scale_fill_manual(values = cbPalette) +
    scale_y_log10(breaks = c(0.0001, 0.001, 0.01, 0.1, 1, 10), labels = function(i) paste(i)) +
    theme_figure(text_size = 10) + labs(y = "calibrated d15 difference with one vs. two step [permil]")) %>% 
ggplotly()
```


### Error estimation
Using the maximum standard deviation of all isotopic standards and controls to make an estimate of the absolute error in the isotopic measurement. 
```{r}
# delta error
d_err <- one_step %>% filter(category %in% c("IAEA-NO3", "USGS-34") | grepl("STD nitrate", name)) %>% 
  group_by(category) %>% 
  iso_summarize_data_table(d15.ocal, amount.nmol_pred) 
d_err
d15.err <- max(d_err$`d15.ocal sd`, na.rm=T)

# add error estimates to sample info
df_final <-one_step %>% 
  select(analysis, run_number, pH, volume, area,name, category, dilution, col_NOx,col_NH4, conc_uM, conc_uM_err, yield, d15.ocal)
df_final <- df_final %>% 
  mutate(
    d15N = d15.ocal,
    d15N_err = d15.err) 

df_final
```

### Check data
```{r "Visualize Calibrated Nitrogen Isotopic Data", plotly = TRUE}
#visualization of calibrated N isotopic data
N_plot <- df_final %>%
	filter(category != "excluded") %>% #remove excluded samples and standards from plot
	ggplot(
		aes(x = run_number, y = d15N, color = category, label = name)) +
			geom_point(size = 4) +
		  geom_errorbar(aes(ymax = d15N + d15N_err, ymin = d15N - d15N_err)) +
			theme_figure() +
	scale_color_manual(values = cbPalette) +
			xlab("Run Number")  +
	guides(color = guide_legend(title = NULL))

#plot interactive version to see individual sample names from plot above
N_plot_interactive <- ggplotly(N_plot + theme_bw()) %>% 
	layout(xaxis = list(title = plotly::TeX("Run  Number")), yaxis = list(title = plotly::TeX("\\delta^{15}N"))) %>% 
	config(mathjax = 'cdn')

N_plot_interactive
```


## Summary
```{r}
df_final <- df_final %>%
  mutate(
    conc = conc_uM,
    conc_err = conc_uM_err,
    compound = "NO3+Norg"
  )  

df_final %>% group_by(category, name) %>% 
  iso_summarize_data_table(cutoff = 1, d15N, d15N_err, conc, conc_err)
```

## Export Sample Data
```{r, include=FALSE}
df_final %>%
  filter(category == "Oman") %>%
  select(analysis, run_number, volume, area, name, category, dilution, col_NOx, col_NH4, pH, conc, conc_err, d15N, d15N_err, compound) %>%
write.csv(file = file.path("data", "calibrated", "run10_calib_data_export.csv"))

df_final %>% 
 filter(category == "Oman") %>% 
  saveRDS(file = file.path("data", "calibrated", "df_final_run10"))
# saveRDS
``` 


## Export standards data
```{r, include=FALSE}
df_final %>%
  filter(category %in% c("USGS-34", "IAEA-NO3", "USGS-40", "USGS-65")) %>%
  select(analysis, run_number, volume, area, name, category, conc, conc_err, d15N, d15N_err, compound) %>%
write.csv(file = file.path("data", "calibrated", "run10_stds_calib_data_export.csv"))

df_final %>% 
 filter(category %in% c("USGS-34", "IAEA-NO3")) %>%
  saveRDS(file = file.path("data", "calibrated", "df_stds_run10"))
# saveRDS
```

